Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
edits to 2.3

The purpose of the recognition model is to accelerate the search over
programs.  It does this by learning to predict which programs both 
assign high likelihood to a task, and at the same time 
have high probability under $(\mathcal{D},\theta)$.

@@

We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]||q(p|x) \right) \right]$,
equivalently maximizing
\begin{equation*}
  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
\end{equation*}
where the expectation is taken over tasks. One could take this expectation
over the observed empirical distribution of tasks,
like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
We found it useful to maximize both an autoencoder-style objective (written $\mathcal{L}_{\text{AE}}$) and a Helmholtz-style objective ($\mathcal{L}_{\text{HM}}$), giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
\begin{align}
\mathcal{L}_{\text{RM}}& = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}\\
\mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\\
\mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p|\mathcal{D},\theta \right]}\log q(p|x)\right]\nonumber
\end{align}
The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity $q$.
Once we have bootstrapped a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$,
and train $q$ on those samples.

@@

The architecture of $q$ depends upon the domain.
For domains with sequential structure like strings and lists,
we use an RNN to encode each input/output example,
and for domains with fixed-dimensionality inputs/outputs we use a multilayer perceptron.
We MaxPool across all of the input/output examples. See Supplement for details.

>>>>

The purpose of the recognition model is to accelerate the search for
programs.  It does this by learning to predict programs which are probable
under $(\mathcal{D},\theta)$ while also assigning high likelihood for a task
according to $\probability[x|p]$.

@@

We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
or equivalently maximizing
\begin{equation*}
  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
\end{equation*}
where the expectation is taken over tasks. One could take this expectation
over the observed empirical distribution of tasks,
like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
\begin{align}
\mathcal{L}_{\text{RM}}& = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}\\
\mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\\
\mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p|\mathcal{D},\theta \right]}\log q(p|x)\right]\nonumber
\end{align}
The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

@@

The architecture of $q$ depends upon the domain.
For domains with sequential structure like strings and lists,
we use an RNN to encode each input/output example.
For domains with fixed-dimensionality inputs/outputs we use a multilayer perceptron.
We MaxPool across all of the input/output examples. See Supplement for details.
